Modern type-safe embedding of attribute grammars

### 2018-11-21 doc/Paper.tex
Update PDF to 787cbdae025f631eb54b7ca7c6e3787074bed149

  Attribute Grammars (AGs) are a declarative formalism that was proposed by
  Knuth~\cite{Knuth68} in the late 60s and allows the implementation and
  reasoning about programs in a modular and convenient way. A concrete AG
  relies on a context-free grammar to define the syntax of a language, and on
  attributes associated to

>>>>

  Attribute Grammars (AGs) are a declarative formalism that was proposed by
  Knuth~\cite{Knuth68} in the late 60s and allows the implementation and
  reasoning about programs in a modular and convenient way. A concrete AG
  relies on a context-free grammar to define the syntax of a language, and on
  attributes associated to the productions of the grammar to define the
  semantics of that language.
  %while adding \emph{attributes} to it so that it is also possible to define
  %its semantics.
  AGs have been used in practice to specify real programming languages, like
  for example Haskell \cite{DijkstraFS09}, as well as powerful pretty printing
  algorithms \cite{SPS99}, deforestation techniques \cite{joao07pepm} and
  powerful type systems \cite{MiddelkoopDS10}.

  When programming with AGs, modularity is achieved due to the possibility of
  defining and using different aspects of computations as separate attributes.
  Attributes are distinct computation units, typically quite simple and modular,
  that can be combined into elaborated solutions to complex programming
  problems. They can also be analyzed, debugged and maintained independently
  which eases program development and evolution.

  AGs have proven to be particularly useful to specify computations over
  trees: given one tree, several AG systems such
  as~\cite{syngen,uuag,lrc,silver} take specifications of which values, or
  attributes, need to be computed on the tree and perform these computations.
  The design and coding efforts put into the creation, improvement and
  maintenance of these AG systems, however, is tremendous, which often is an
  obstacle to achieving the success they deserve.

  An increasingly popular alternative approach to the use of AGs relies on
  embedding them as first class citizens of general purpose programming
  languages~\cite{Oege00,DBLP:conf/sblp/MartinsFS13,erlangAGs,kiama,doaitse09icfp,balestrieri}.
  This avoids the burden of implementing a totally new language and associated
  system by hosting it in state-of-the-art programming languages. Following this
  approach one then exploits the modern constructions and infrastructure that
  are already provided by those languages and focuses on the particularities of
  the domain specific language being developed.

  Functional zipper~\cite{huet1997zipper} is a powerful abstraction which greatly
  simplifies the implementation of traversal algorithms performing a lot of
  local updates. Functional zippers have successfully been applied to constuct
  an attribute grammar embedding in Haskell~\cite{DBLP:conf/sblp/MartinsFS13,MARTINS20162}. Despite its elegance,
  this solution had a major drawback which prevented its use in real-world
  applications: attributes were not cached, but rather repeatedly recomputed
  which severely hurt performance. Recently, this flaw has been
  eliminated~\cite{FERNANDES2018} and replaced with a different one: the approach became
  intrusive, i.e. to benefit from the embedding user-defined data structures
  have to be adjusted.

  In this paper we present an alternative mechanism to cache attributes based on
  a self-organising infinite grid. This graph is laid on top of the user-defined
  algebraic data type and mirrors its structure. The used-defined data type
  remains untouched. The embedding is then based on two coherent zippers (rather
  than one) traversing the data structures in parallel. On top of being
  non-intrusive our solution is completely type-safe. Modern Haskell extensions
  such as \texttt{ConstraintKinds} allow us to propagate constraints down in the
  ADT completely eliminating run-time type casts present in the previous
  versions. Another side benefit of using modern Haskell is a cleaner syntax
  with less code being generated with Template Haskell.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% As per the LLNCS guidelines
\bibliographystyle{splncs04}
\bibliography{References}

\end{document}
